\vspace{-2cm}A generative model for urban firm networks

### 2022-03-28 Response_to_reviewers.tex

% redo map with open resources (see editor request) https://www.naturalearthdata.com/downloads/110m-cultural-vectors/

\textit{Maps should use only open resources.}

\medskip

$\rightarrow$ The map in Figure 2 was redrawn using countries from Natural Earth Data (public domain).

@@

	$\rightarrow$ Additional statistical models extended to account for zeros were fitted and commented. This aspect was better explained.
	%(JR) Hurdle and zero inflated models\comment{(JR) Discuss choice/possibly test other stat models}

	\medskip

%1.3 -

	\item \textit{Given that the model is generative, the temporal aspect of the model can be an issue. Indeed, if the number of iterations (around 5000) needed to generate the empirical network correspond to a long time period (for example decades), the economic size and sectorial proximity are expected to change, probably implying a substantially different network, if these changes were accounted for in the model. Can the authors mention (and may be discuss) this issue?}

	$\rightarrow$ The way the model captures time steps and deals with time granularity was indeed not discussed well in the previous version of the paper. Effective time simulated by the model can be set by modifying both $t_f$ (number of steps) and $w_0$ parameters. The model assumes that context remains constant, so that medium-length periods can be captured. This was better explained in the model description.
	% develop - (JR) total time is linked to $w_0$ - discuss this

@@

	$\rightarrow$ The purpose of both the statistical analysis and the generative model is not prediction. Statistical analysis is a preliminary exploration to determine potential processes to be included in the generative model, while the latest is an explanatory model, i.e. aimed at explaining the role of different factors in the growth of the urban network (such as the role of urban hierarchy) and test hypothesis in a virtual laboratory. It is also used to prototype policy applications, and in that sense is potentially a decision-making model. A discussion on model function and the precise positioning of this model was added.
	%\todo{(JR) this is not a prediction model - possibly give values of abs error - discuss more model function}

	\medskip

	\item \textit{2) The authors do not show how well the model reproduces the metropolisation and the communities of the empirical network. If the model does not perform well at reproducing these properties of the network, it can mean that the predictors they used are not relevant to such features. This would imply that the model is of limited relevancy to study the impact of crises on these structural features.}

	$\rightarrow$ We added the computation of empirical macroscopic indicators (internationalisation and metropolisation), and how the model gets close to these through the exploration. The calibration at the link level (mean square errors) remains the most relevant for discussion - this was further discussed.
	%\todo{(JR) compute more empirical indicators and compare  with generated networks}

@@

\begin{enumerate}

%1.

	\item \textit{Clarify the paper's aims. The part about border closures is interesting but seems almost tacked-on. Is the paper about 'what if' borders closed (e.g. COVID, Brexit), or is it about the communities that exist and the relative hierarchies within the network. To be honest, I don't know if I could pick one after reading it. The conclusion claims that "this paper [is] aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset". This should be reoriented toward a more theoretical framing, and speak to the network geographic ties behind it.}

	$\rightarrow$ The paper positioning and goal were reframed. The purpose is to introduce a generative model to understand how different geographical and economic factors influence the growth of urban firm networks, with both potentialities for theoretical explanations (through exploration and calibration) and practical applications (the latest being a proof-of-concept in this work).
	%\todo{(All) paper positioning}

@@

	$\rightarrow$ \todo{(All) Literature on geo-economic factors has been enhanced - processes included}

	\medskip

%5)
	\item \textit{The application of economic shocks appears to be an additional part of the model that is not properly discussed again in the literature review. I wonder if this is actually a different paper? I felt there were lots of interesting aspects to the research - and thinking about how to divide out the components to allow better discussion for each would benefit the research in that you could dive alot deeper into specific discussions.}

	$\rightarrow$ Echoing a remark by the second referee, the model application is not the main focus of the paper, but a proof-of-concept illustration for possible developments. This was clarified in paper positioning.
	%\todo{(All) justify proof of concept - add literature}

### 2022-03-28 main.tex
Update on Overleaf.

\section*{Abstract}
The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level including several complementary processes: the economic size of urban areas at origin and destination of links, industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas. An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We calibrate the model on the European network, outperforming statistical models showing a strong role of path-dependency. The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lockdowns of countries. This is illustrated with an application on stylised scenarios.

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}. In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}. Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

In this paper we choose to focus on interurban interactions between firms considering the preferential attachment paradigm in a network \cite{BarabasiAlbert1999} and spatial interaction model properties % add a paragraph with the most important literature on spatial interaction models and preferential attachment in network and how it can be linked to firm networks.
Firm networks remains an important aspect of network and system of cities theory \cite{RozenblatNeal2021}. They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.

Following this literature the present paper positions itself at the intersection of global networks theoretical framework, spatial interaction models and preferential attachment. It contributes to question and capture geographical and economic processes as internationalisation, metropolisation, regionalisation and specialisation of cities in terms of firm networks within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}. In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}.

@@

The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and proceed to an empirical analysis of the network used as a case study. We then proceed to a sensitivity analysis and exploration of the model on synthetic data. The model is calibrated on empirical European firm ownership data. We finally discuss implications of our results and possible extensions.

@@

\subsubsection*{Rationale}

The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several processes driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (confirmed relevant by the fixed effect regression described below); (iii) city economic size, which presents agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages (influence of already existing links in the past). This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes in the growth of the urban network and to test diverse hypothesis \cite{conte2012manifesto}. It also tends to be a decision-making model as we explore potential policy applications. Our focus on explanation rather than prediction can be understood in a context of a broad diversity of possible model functions \cite{varenne2018models}.

@@

\begin{equation}
p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
\end{equation}
where $E  =  \sum_k E_k$, $W  = \sum_{i,j} w_{ij}$ weights of previous links, $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$ taken as a cosine similarity, $d_{ij}$ is the Euclidian distance, and $c_{ij}$ is a socio-cultural distance. In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$. The model is stopped either when a final time is reached or when the cumulated volume of links reaches a target quantity. This formulation can be considered both as an economic utility model and a generalisation of spatial interaction models.

Note that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions. The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting. In this simple version of the model we assume an absence of evolution of economic size, which means that the adjusted parameters are valid on a restricted time period.

@@

In synthetic experiments, the network is generated from an initial complete network with dummy links of weight 1 until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.

@@

Secondly, we consider the geographical process of metropolisation. The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}. Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}. Metropolisation gradients have been already explored in literature in city-systems considering several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}. We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.

@@

We base our study on empirical data from \emph{ADAMDEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe. AMADEUS indicates the exact geolocation, turnover and activity codes for owners and owned companies in all sectors in Europe. The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet. A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}. \cite{bajgar2020coverage} find that the coverage of the \emph{ORBIS} database on OECD countries is a representative sample, but with a better performance on larger firms. As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
% \comment[JR]{To what database is Orbis compared to in this study - what is the confidence on its exhaustivity}

@@

%%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{figures/Fig2.png}
    \end{center}
    \vspace{2cm}
    \caption{{\bf Map of network communities obtained by modularity maximisation.} We show communities with the higher modularity, obtained with the Louvain algorithm. The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to. Urban Areas in communities of size smaller than 3 were not coloured for readability.}
    \label{fig:fig2}
\end{figure}
%%%%%%%%%%%%%%

% TODO redo map with non licensed data for countries: https://www.naturalearthdata.com/downloads/110m-cultural-vectors/

@@

We then apply the model to a real system of cities by calibrating it on the European ownership network. Model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process. The objective functions for calibration are the average mean squared error on logarithms of weights given by $\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2$ if $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error $\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)$. \cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models. Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model. Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples. We use a population of 200 individuals and run the algorithm for 50,000 generations, on 1000 islands in parallel.

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    \end{center}
    \vspace{2cm}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrizations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%

@@

When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$. This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency. The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models). The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe. Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data.

% TODO recall not a predictive model

@@

This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMole platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset. The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space. Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions. Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values). This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.

The calibration of model suggests its potential application to policy issues and can therefore have practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in details are: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text). % ~\ref{S1Text} ref not working

>>>>

\section*{Abstract}
The emergence of interconnected urban networks is a crucial feature of globalisation processes. Understanding the drivers behind the growth of such networks -- in particular urban firm networks --, is essential for the economic resilience of urban systems. We introduce in this paper a generative model for firm networks at the urban area level, which considers the following features: the economic size of the areas representing the origin and destination nodes, the industrial sector proximity between firms, the strength of links from the past, as well as the geographical and socio-cultural distance between urban areas. An empirical network analysis on European urban firm ownership data confirms the relevance of each of these factors on new link creation. We then simulate network growth on synthetic systems of cities, unveiling stylized facts such as a transition from a local to a global regime or a maximal integration achieved at an intermediate interaction range. We calibrate the model on the European network, outperforming statistical models showing a strong role of path-dependency. The model opens up potential applications to formulate public policies recommendations to deal with exogenous shocks such as economic crisis or potential lockdowns of countries. This is illustrated with an application on stylised scenarios.

@@

Networks constitute the new social morphology of our societies, since the intensification of globalisation of the economy in the late 20th century \cite{castells2000networksociety}. A world-city system has since then emerged \cite{taylor2001specification}, characterized by highly interconnected urban centres at the global level \cite{sassen1991}. These interactions can be of different nature and determined by economic, socio-cultural, or geopolitical proximity inducing spatial and non-spatial patterns \cite{martinus2018global}. Interactions of economic nature between transnational firms \cite{derudder2018central}, are together with international trade patterns among the most representative of the current economic trends \cite{taylor2001specification}. Understanding the drivers of growth and geographical properties of such an urban network, can be used to foster innovation in specific cities and to shape public policies for local industrial clusters \cite{turkina2016structure}. In addition, strengthening these city-interactions can be crucial for revitalising certain geographical areas \cite{Clarke2018}. This can be achieved through strategic development of certain industrial sectors for global integration of regions \cite{dawley2019creating}, cities \cite{gluckler2016relational} and countries \cite{martinus2019brokerage}. Exploring city/firm interactions \cite{storme2019introducing} and the position of cities within these networks \cite{gluckler2016relational} can also permit to infer the consequences of future unforeseen exogenous shocks such as an economic crisis, a post-Brexit scenario for the European Union or a single market collapse in the post-COVID19 future.

Network models to understand processes underlying the emergence and dynamics of such networks have been proposed in the literature, nevertheless, these are not extensive and do not explicitly model the city level \cite{taylor2001specification}. At regional scale, generative urban network models combining geographical factors with network topological properties have been proposed \cite{dai2016simulating}. At country level, trade networks have been extensively studied from the complex networks perspective \cite{garlaschelli2005structure}, but these are generally not explored at a more granular level due to the lack of city-level data. Within regional science, input-output models are used as generative urban models of flows between geographical areas \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect} at different geographical scales.

%aca: change

% add 3 mains contributions to literature of this paper 1. approach by cities and not regions, approach by cities and not firm to firm links only 2. manière dont on modélise le poids avec le cobb-douglas pas avec le GDP, mais avec les % owernship links - vérifier ce que les économiustes utilisent pour modéliser les flux 3. modèle génératif et dynamique 
In this paper we choose to focus on interurban interactions between firms. Firm networks remains an central aspect of network and system of cities theory \cite{RozenblatNeal2021} from the economic perspective. They have already been investigated in complexity science at city level to unveil agglomeration economies and competitiveness of cities \cite{Rozenblat2010}, regionalisation patterns \cite{Rozenblatetal2017} or to model hierarchy and specialisation \cite{BidaRozenblat2020}. Moreover economic complexity indicators have been developed in this sense to characterise the production structure and specialisation of different countries due to economic activity, including those of trade firms \cite{HaussmanHidalgo2009,OCleary2016}.

Following this literature the present paper contributes to question and capture geographical and economic processes as internationalisation and metropolisation in terms of firm networks within a generative model. The specific question we tackle is the complementarity of different processes to generate such urban networks. The originality of the present paper is to examine these issues within the system of cities \cite{berry1964cities} by following previous literature the field of urban firm networks \cite{Rozenblatetal2017}. In this approach the position and dynamics of cities in the socio-economic world-wide space can be considered by their interactions with other cities \cite{pumain2018evolutionary}. We examine the interactions of European cities within firm linkages defined by corporate ownership links. Since transnational firms structure is one of the determinant of the global economic space, it provides one accurate proxy to unveil geo-economic structures \cite{2020Zdanowska}. 
% regionalisation and specialisation of cities
% aca: to here

%The use of network models to understand processes underlying the emergence and dynamics of such networks has already been tackled in the literature, but still not extensively and never explicitly at city level \cite{taylor2001specification}. In fact generative urban network models combining geographical factors with network topological properties exist at the regional scale {\cite{dai2016simulating}}. Trade networks have been extensively studied from the complex networks perspective, but these are generally driven at the country-level \cite{garlaschelli2005structure} because of a lack of city-level data. Input-output models, mostly used in regional science to represent flows between geographical areas, are considered in some cases as a type of urban generative model \cite{jin1993generation}. In this stream of research, spatial interaction models \cite{dennett2013multilevel} can form the basis of urban network models for different types of flows \cite{dai2016generative} or physical infrastructure \cite{raimbault2018indirect}.

% Positioning a preciser

@@

%aca: review and extend this
The rest of this paper is organized as follows: we firstly describe the generative model for urban networks and then present an empirical analysis of the network used as a case study. We then perform a sensitivity analysis and exploration of the model parameters using synthetic data. The model is then calibrated on empirical European firm ownership data. We finally discuss implications of our results and possible extensions.

@@

\subsubsection*{Rationale}
%aca: change
The model is expected to capture a single macroscopic urban scale, i.e. the level of the urban system where basic entities are cities. Links are induced by underlying firms but these are not explicit. Cities are defined by their economic size, but also by an industrial sector profile. As the authors of \cite{martinus2018global} point out, a combination of several distances may play a role in establishing linkages. %aca: to here
Geographical distance, industrial proximity and complementarity will typically be equally significant \cite{cottineau2020nested}. Therefore, we combine several features %processes 
driving network growth: (i) geographical proximity (in our case the crow-fly distance, but it can be generalised to any type of accessibility); (ii) geopolitical proximity, which captures the higher propensity to establish links within existing submarkets within the European common market (this is confirmed to be  relevant by the fixed effect regression described below); (iii) city economic size, which is known to give rise to agglomeration effects; (iv) industrial proximity in terms of firm sector composition; and (v) previous linkages, which capture the influence of existing linkages created in the past. This last point is crucial to capture accumulated effects and include path-dependencies, justifying further the use of a simulation model instead of a more simple statistical model.

It is worth mentioning that our model is not aimed to be predictive, but rather designed as an explanatory model in order to disentangle the influence of diverse processes and drivers in the growth of the urban network, and to test different hypotheses \cite{conte2012manifesto,varenne2018models}. 
The model can also be used a decision-making model, where different scenarios for potential policy applications can be explored. 
%It also tends to be a decision-making model as we explore potential policy applications. Our focus on explanation rather than prediction can be understood in a context of a broad diversity of possible model functions \cite{varenne2018models}.

@@

\begin{equation}
p_{ij} \propto \left(\frac{E_{i}}{E}\right)^{\gamma_O} \cdot \left(\frac{E_{j}}{E}\right)^{\gamma_D} \cdot \left(\frac{w_{ij}}{W}\right)^{\gamma_W} \cdot s\left(S_{ik},S_{jk}\right)^{\gamma_S} \cdot \exp \left(- d_{ij} / d_0\right) \cdot \exp \left(- c_{ij} / c_0\right)
\end{equation}
where $E  =  \sum_k E_k$; $w_{ij}$ are the weights of previous links, and $W  = \sum_{i,j} w_{ij}$; $s(S_{ik},S_{jk})$ is a similarity function between activity sectors $S_{ik}$ and $S_{jk}$, taken to be the cosine similarity; $d_{ij}$ is the Euclidian distance; and $c_{ij}$ is a socio-cultural distance. In the case of an already existing link between two areas, the weight of the latter is incremented by a unit volume $w_0$. The model is stopped either when a final time $t_f$ is reached or when the cumulative volume of links reaches a target quantity. Therefore, the duration of the time period simulated can be adjusted by changing either $t_f$ or $w_0$ parameter, given a total volume.

This model formulation can be considered both as an economic utility model and a generalisation of spatial interaction models. We must furthermore precise that: (i) we consider an asymmetric influence of sizes, assuming that link directions are important (similarly the similarity function $s$ may be taken as asymmetric); (ii) the influence of previous links is similar to a preferential attachment process; (iii) we do not renormalise the exponents to 1 to ensure to include convex functions. The ``geopolitical/sociocultural'' distance remains abstract and should be parametrised (see setup below) and estimated in the data-driven setting. In this simple version of the model we assume an absence of evolution of economic size or sectorial proximity, which means that the adjusted parameters are valid on a restricted time period.

@@

In synthetic experiments, the network is generated from an initial complete network with dummy links of weight $1$ until reaching $t_f=1500$, with a unit link volume of $w_0 = 1$.
% missing links are not constrained

@@

Secondly, we consider the geographical process of metropolisation. The later can be defined as a concentration of population and major economic decision functions in metropolises, in particular at the world-system level \cite{sassen1991global}. Metropolises playing the role of gateways of globalisation, are competing between each other to attract the most qualified and dynamic activities \cite{RozenblatPumain2007}. Metropolisation gradients have been linked in the literature on systems of cities with several economic and political aspects \cite{RozenblatPumain2018} \cite{PumainRozenblat2019} \cite{Zdanowskaetal2020}. We introduce in this paper a metropolisation indicator defined as the correlation between the weighted degree and the GDP of the city, in order to capture the effect of large cities or metropolises of regional importance attracting the biggest amount of links and involved in high interactions with other cities.

@@

We base our study on empirical data from the \emph{AMADEUS} dataset produced by the Bureau Van Dijk (BVD) on transnational firms in Europe. \emph{AMADEUS} provides a precise location, turnover and activity codes for owners and owned companies in all sectors in Europe. The Bureau Van Dick gathers, homogenizes and completes the information on these firms from national sources, publications of consultancies and audit firms such as KPMG, Deloitte, PricewaterhouseCoopers (PwC), Ernst \& Young (EY), Mazars, and those of other financial organisations such as Moody's, Fitch, Standard \& Poor's, Thomson Reuters or Duns and Bradstreet. A representativity study shows that \emph{ORBIS} --the Bureau Van Dijk's dataset for firms at the worldwide scale--, captures around 10 percent of existing foreign direct investments in 2011 in Poland \cite{Zdanowska2018}. \cite{bajgar2020coverage} find that the coverage of the \emph{ORBIS} database on OECD countries is a representative sample, but with a better performance on larger firms. As we aggregate company turnovers on urban areas in our analysis, we thus do not expect much representativity bias in the inter-urban network.
% To what database is Orbis compared to in this study - what is the confidence on its exhaustivity

@@

%%%%%%%%%%%%%%
\begin{figure}
    \begin{center}
    \includegraphics[width=\linewidth]{figures/Fig2.png}
    \end{center}
    \vspace{2cm}
    \caption{{\bf Map of network communities obtained by modularity maximisation.} We show communities with the higher modularity, obtained with the Louvain algorithm. The area of circles gives the turnover within the Functional Urban Area, and colour the community it belongs to. Urban Areas in communities of size smaller than 3 were not coloured for readability. Made with Natural Earth Data \cite{naturalearth}.}
    \label{fig:fig2}
\end{figure}
%%%%%%%%%%%%%%

Model macroscopic indicators of internationalisation and metropolisation can also be computed on the empirical network. Internationalisation corresponds to the modularity of countries in the network, which was computed above as $I = 0.32$. Metropolisation computed with turnover has a very high empirical value $M = 0.961$ (95 \% confidence interval $\left[ 0.954 ; 0.967\right]$). A variation of this indicator computed with GDP from the GHSL database is not as high but still a relatively high correlation ($0.776 \left[ 0.736 ; 0.811\right]$). In practice, this European urban network is thus highly internationalised and metropolised, following what can be expected from the empirical literature on urban networks.

@@

We then apply the model to a real system of cities by calibrating it on the European ownership network. Model setup is done following the real setup described above. The number of time steps $t_f$ is left as an additional parameter, which in a sense determines the granularity of the cumulative network generation process. The model is fitted at the microscopic level on link structure: the objective functions for calibration are the average mean squared error on logarithms of weights given by

\begin{equation}
\varepsilon_L = \frac{1}{N^2} \sum_{i,j} \left(\log w_{ij} - \log \hat{w}_{ij} \right)^2
\end{equation}

where $\hat{w}_{ij}$ are the simulated weights, and the logarithm of the mean squared error

\begin{equation}
\varepsilon_M = \log\left(\frac{1}{N^2} \sum_{i,j} \left(w_{ij} - \hat{w}_{ij}\right)^2 \right)
\end{equation}

\cite{raimbault2018indirect} showed that these two are complementary to calibrate urban systems models, capturing the overall performance but also how the model behaves on medium-sized and small links. A calibration on macroscopic indicators of metropolisation and internationalisation would not bring much knowledge on model processes, since several points from the grid experiment above - not designed for this - already provide a good performance (19 parameter points out of 6534 are simultaneously below a relative error of 10\% on internationalisation and below 10\% on metropolisation).
% TODO in progress: test macro calib ; rerun explo metropolisation? -> ok

Calibration is done using a genetic NSGA2 algorithm \cite{deb2002fast}, which is particularly suited for such a bi-objective calibration of a simulation model. Stochasticity is taken into account by adding the estimation confidence as an additional objective, and we filter the final population to have at least 20 samples. We use a population of 200 individuals and run the algorithm for 50,000 generations, on 1000 islands in parallel.

%%%%%%%%%%%%%%
\begin{figure}
	\begin{center}
    \includegraphics[width=0.75\linewidth]{figures/Fig8.png}
    \end{center}
    \vspace{2cm}
	\caption{\textbf{Model calibration results.} Pareto front obtained for the bi-objective model calibration for $\varepsilon_L$ and $\varepsilon_M$ objectives. $\varepsilon_L$ can be compared to the MSE of statistical models fitted above, and the generative model outperform these for several parametrisations (lowest MSE 5.33 for statistical models). Point size gives the number of stochastic samples and color the value of $\gamma_O$.\label{fig:fig8}}
\end{figure}
%%%%%%%%%%%%%%

@@

When considering optimal points with $\varepsilon_L < 5.0$, the adjusted parameters are: long range gravity interactions and socio-economic interactions given by $d_0 = 5972 \pm 2148$ and $c_0 = 113 \pm 33$ (the maximal $c_{ij}$ is 11.8); a stronger influence of destination than origin as $\gamma_O = 1.46 \pm 0.10$ and $\gamma_D = 2.0 \pm 0.05$; a comparable influence of sectors as $\gamma_S = 1.76 \pm 0.21$; a fine time granularity as $t_f = 4951 \pm 124$; and a very strong influence of previous links as $\gamma_W = 6.43 \pm 4.11$. This last point furthermore shows in particular a role of the path-dependency parameter, confirming the relevance of this complex generative model including path-dependency. The interaction of this process with others implies an inversion of the relative importance of origin and destination (on the contrary to statistical models). The fact that calibrated interaction distances are large is consistent with long range economic interactions within Europe. Besides these stylized facts, these calibration results altogether demonstrate the feasibility of model application to real data. As stated previously, the model is not predictive and the relatively high error values would not make it satisfactory in that manner, but are reasonable for its explanatory and application functions.

@@

This paper aimed at presenting a generative model for urban networks defined by interactions between firms based on synthetic data, simulated via the OpenMOLE platform and calibrated on real data on ownership linkages of firms for Europe from the Amadeus dataset. The simulation on a synthetic system of cities provides a broad knowledge on model behavior in its parameter space. Even in such a simple model (close to directly tractable stationary state) the behaviour is highly non-linear in many dimensions. Our study shows how crucial model exploration is to overcome hidden parameters (deactivated mechanisms or default parameter values). This paper emphasis on the fact that exploration of intrinsic dynamics on synthetic data is of great importance and should be driven before an application on real data so as to put aside the geographical effects from model dynamics.

The calibration of model suggests its potential application to policy issues and can therefore have practical applications in the future. For example the effect of exogenous shocks in the socio-economic structure can be tested as we did in the stylized application example above. Such shocks can for example be the upcoming impact of Brexit on the European market or the consequence of a long-term closure of the European Union members' borders due to further pandemic outbreaks on the single market. In this sense the UK economic sectors mostly dependent on foreign capital and generating the highest turnovers in 2018, that would be the most affected and worth examining in details are: extraction of crude petroleum, manufacture of tobacco products and of basic pharmaceutical goods (in terms of out-going links from the UK) ; activities of head offices and holding companies, sale of cars and manufacture of motor vehicles (in terms of in-going links to the UK). We showed in our experiment that although the system is resilient to moderate shocks, strong restrictions have an important impact on internationalization and thus can be expected to have a detrimental effect on UK economy due to these foreign ownership links (see also S1 Text\ref{S1Text}). The impact of such unforeseen economic shocks could also be considered from a multi-level perspective, coupling our work with other approaches such as \cite{Dorretal2022} studying the firm-level impact of the Covid crisis.
